---
author: 'Keanu Rochette '
date: '`r format(Sys.Date())`'
title: Final Project
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning= FALSE)
```

# Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(openxlsx) 
library(gt)
```


################# Data Clean-Up for CTD Data ####################

# Create a function that converts .asc files to .csv
```{}
asc_to_csv <- function(filename){
  
# Specify the file paths
asc_file <- here("Data_manipulation", "raw_data","ctd_data", filename)  

# Read the .asc file
data <- read.table(asc_file, sep = ",")  
colnames(data) <- data[1,]
data <- data[-1,]
data <- clean_names(data)


filename <- str_replace_all(filename, "\\.|[a-z]{3}", "")
filename <- str_trim(filename)

# Write to a .csv file
write.csv(data, file = here("Data_manipulation","raw_data","ctd_data",
                            paste0(filename, ".csv")))

}

```

## Converting the files from .asc to .csv
```{}
asc_to_csv(filename = "km2315_A_1_200m.asc.csv")
asc_to_csv(filename = "km2315_05_500m.asc.csv")
asc_to_csv(filename = "km2315_A_2_200m.asc.csv")
asc_to_csv(filename = "km2315_B_1_250m.asc.csv")
asc_to_csv(filename = "km2315_B_2_250m.asc.csv")
asc_to_csv(filename = "km2315_C_1_250m.asc.csv")
asc_to_csv(filename = "km2315_C_2_250m.asc.csv")
asc_to_csv(filename = "km2315_D_1500m.asc.csv")

#Note: Somehow the file names have a space before the ".csv" and I don't know how 
# to fix that. So I manually removed the spaces. 
# Update: paste0() should do the trick
```


#Load Data
```{r}
CTD_A1 <- read_csv(here("Data_manipulation", "raw_data", "ctd_data", "km2315_A_1_200m.csv"))
CTD_A2 <- read_csv(here("Data_manipulation", "raw_data","ctd_data", "km2315_A_2_200m.csv"))
CTD_B1 <- read_csv(here("Data_manipulation", "raw_data","ctd_data", "km2315_B_1_250m.csv"))
CTD_B2 <- read_csv(here("Data_manipulation", "raw_data","ctd_data", "km2315_B_2_250m.csv"))
CTD_C1 <- read_csv(here("Data_manipulation", "raw_data","ctd_data", "km2315_C_1_250m.csv"))
CTD_C2 <- read_csv(here("Data_manipulation","raw_data","ctd_data", "km2315_C_2_250m.csv"))
CTD_D <- read_csv(here("Data_manipulation", "raw_data","ctd_data", "km2315_D_1500m.csv"))


```

## Only keep the down cast 

```{r}
max_time <- CTD_A1 %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_A1 <- CTD_A1 %>% filter(time_s < max_time)

max_time <- CTD_A2 %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_A2 <- CTD_A2 %>% filter(time_s < max_time)

max_time <- CTD_B1 %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_B1 <- CTD_B1 %>% filter(time_s < max_time)

max_time <- CTD_B2 %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_B2 <- CTD_B2 %>% filter(time_s < max_time)

max_time <- CTD_C1 %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_C1 <- CTD_C1 %>% filter(time_s < max_time)

max_time <- CTD_C2 %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_C2 <- CTD_C2 %>% filter(time_s < max_time)

max_time <- CTD_D %>% filter(pr_dm == max(pr_dm)) %>% pull(time_s)
CTD_D <- CTD_D %>% filter(time_s < max_time)

```

# Create a function to automate the clean up the CTD data
```{r}

CTD_data_cleaner <- function(dataset){

  station_ID <- 
  df <- dataset %>% 
    mutate(temp = (t090c+t190c)/2,
         density = (density00+density11)/2,
         date = as.Date(time_j, origin= as.Date("2023-01-01")),
         oxygen_conc = (sbox0mm_kg+sbox1mm_kg)/2,
         perc_ox_sat = (sbeox0ps+sbeox1ps)/2) %>% 
    rename(pressure = pr_dm, depth = dep_sm,
         descent_rate= dz_dt_m, 
         PAR = par,
         fluorescence = fl_sp) %>% 
    select(time_s, date, pressure, descent_rate, PAR, 
         temp, density, oxygen_conc, perc_ox_sat, fluorescence, c_star_at0, c_star_tr0)  
  
    breaks <- seq(from = 0, to = max(df$pressure), by = 0.5)
    
    df <- df %>%  tibble() %>% 
      mutate(depth_cat = cut(df$pressure, 
        breaks= breaks,
        labels=paste(head(breaks, -1)),
        right = F)) %>% 
      relocate(depth_cat, .before = time_s) %>% 
    group_by(depth_cat) %>% 
    summarise_all(mean) %>% ungroup()
    
    return(df)
  
}
```

# Apply the function to the numerous files and clean up the data 
```{r}
CTD_A1 <- CTD_data_cleaner(CTD_A1)
CTD_A1 <- CTD_A1 %>% mutate(station = "A1") 

CTD_A2 <- CTD_data_cleaner(CTD_A2)
CTD_A2<-CTD_A2 %>% mutate(station = "A2") 

CTD_B1 <- CTD_data_cleaner(CTD_B1)
CTD_B1 <-CTD_B1 %>% mutate(station = "B1") 

CTD_B2 <- CTD_data_cleaner(CTD_B2)
CTD_B2<-CTD_B2 %>% mutate(station = "B2") 

CTD_C1 <- CTD_data_cleaner(CTD_C1)
CTD_C1 <-CTD_C1 %>% mutate(station = "C1") 

CTD_C2 <- CTD_data_cleaner(CTD_C2)
CTD_C2<- CTD_C2 %>% mutate(station = "C2")

CTD_D <- CTD_data_cleaner(CTD_D)
CTD_D <- CTD_D %>% mutate(station = "Deep")

```


## Saving the files for Devin

```{r}
write_csv(CTD_A1, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_A_1_200m.csv"))

write_csv(CTD_A2, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_A_2_200m.csv"))

write_csv(CTD_B1, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_B_1_250m.csv"))

write_csv(CTD_B2, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_B_2_250m.csv"))

write_csv(CTD_C1, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_C_1_250m.csv"))

write_csv(CTD_C2, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_C_2_250m.csv"))

write_csv(CTD_D, file = here("Data_manipulation", "raw_data", "ctd_data", "Devin", "km2315_D_1500m.csv"))
```


# Combining all the CTD data into one file 
```{r}
CTD <- rbind(CTD_A1 ,CTD_A2, CTD_B1,CTD_B2, CTD_C1, CTD_C2, CTD_D)
```


# Final Data clean up and saving the file as a new .csv
```{r}
CTD <- CTD %>% group_by(depth_cat, station) %>% 
  summarise_all(mean) %>% ungroup() 
  
write_csv(CTD, here("Oceanographic_data", "data", "CTD.csv"))
  
```


################# Data Clean-Up for CTD Data ####################

# Load the Data
```{r}
fcm <- read_csv(here("Oceanographic_data", "data","raw_data","fcm_data", "km2315_fcm_data.csv"), locale=locale(encoding="latin1"), skip = 2)

site_depth <- read_csv(here("Oceanographic_data", "data","raw_data","fcm_data", "wells_dataframe.csv"))

#only keep the 1/4L(V)
# relevents = (include - hoechst junk - chlorophyll junk)
```

# Clean the column names of the files

```{r}
fcm <- fcm %>% clean_names() 
site_depth<- site_depth %>% rename(tube_name = well_num)
```

# Renames the columns and only keeping the columns of interest
```{r}
fcm <- fcm %>% 
  mutate(real_event = include_events_i1_4l_v- hoechst_junk_events_i1_4l_v- chlorophyll_junk_events_i1_4l_v) %>% 
  select(1,c(166:181), -h1_ur_events_i1_4l_v, -h1_ul_events_i1_4l_v,
         -h2_ll_events_i1_4l_v, -h2_lr_events_i1_4l_v,
         -hoechst_junk_events_i1_4l_v, -chlorophyll_junk_events_i1_4l_v
         ) %>% 
  rename(all_events = all_events_events_i1_4l_v,
         include_events = include_events_i1_4l_v,
         low_chl_events = low_chlorophyll_events_i1_4l_v,
         autotr= autotrophic_hi_chl_events_i1_4l_v,
         small_autotr_cyano = small_autotrophs_cyanobacteria_events_i1_4l_v,
         large_autotr_pico_euks = large_autotrophs_pico_euks_events_i1_4l_v,
         heterobact = heterotrophic_bacteria_events_i1_4l_v,
         low_pe_cyano = low_pe_cyanos_events_i1_4l_v,
         hi_pe_cyano = high_pe_cyanos_events_i1_4l_v,
         true_autotr = true_autotrophs_events_i1_4l_v)
```

#Merging the data sets by sample ID so we have a comprehensive dataframe

```{r}
fcm <- left_join(fcm, site_depth, by= "tube_name")
fcm <- fcm %>%  relocate(site_depth, .after = tube_name)
```

# Separating the site name and cast number for later analysis
```{r}
fcm <- fcm %>% separate(site_depth, c("site", "depth")) %>% 
  separate(site, 
           into = c("site", "cast"), 
           sep = "(?<=[A-Za-z])(?=[0-9])") %>% 
  filter(site != "stain", site != "blank")
```

# Calculating the relative abundance of microbes 
based on site and depth
```{r}
fcm <- fcm  %>% 
  group_by(site,depth) %>% 
  summarize(hbac_rel_abun = heterobact/include_events,
            pico_euk_rel_abun = large_autotr_pico_euks/include_events,
            syn_rel_abun = hi_pe_cyano/include_events,
            pro_rel_abun = low_pe_cyano/include_events)

```

# Create a new .csv file with clean data
```{r}
write_csv(fcm, here("Oceanographic_data", "data", "fcm.csv"))
```

################# Data Clean-Up for CTD and FCM Data ####################


# Load the data
```{r}
CTD <- read_csv(here("Oceanographic_data", "data", "CTD.csv"))
fcm <- read_csv(here("Oceanographic_data", "data", "fcm.csv"))
```

# Pivoting the dataframe and giving prettier names to the microbes
```{r}
fcm <- fcm %>% 
  pivot_longer(cols = c(hbac_rel_abun:pro_rel_abun),
               names_to = "microbe", 
               values_to = "rel_abun") %>% 
  mutate(microbe = ifelse(microbe == "hbac_rel_abun", "Heterotrophic Bacteria",
                          ifelse(microbe == "pico_euk_rel_abun", "Pico Eukaryote", ifelse(microbe == "syn_rel_abun", "Synechococcus", "Prochlorococcus"))))

         
```

# Only selecting the variable of interest in CTD and pivoting the dataframe
```{r}
CTD <- CTD %>% select(-time_s, -date, -descent_rate, - density) %>% 
   rename(depth = depth_cat, 
            Pressure = pressure,
            `Temperature (°C)` = temp,
            `Oxygen (umol/kg)`= oxygen_conc,
            `Oxygen (% saturation)`=perc_ox_sat,
            Fluorescence = fluorescence
             ) %>% 
  pivot_longer(cols= c(Pressure:Fluorescence),
               names_to = "parameter",
               values_to = "values")
```


```{r}
## CTD Graph
CTD %>% 
      mutate(station = as.factor(station)) %>% 
      filter(station == "A") %>% 
  ggplot(aes(x= depth_cat)) +
  geom_point(aes(y = temp), color = "#FF9800") +
          geom_line(aes(y = temp), color = "#FF9800")+
  coord_flip()+
  labs( title= "Oceanographic Profile",
            x = "Depth (m)",
            y = "Variables") +
      coord_flip()+
      theme_bw() + 
      theme(plot.title = element_text(size=14, face = "bold"), 
            plot.subtitle = element_text(size=12),
            axis.title = element_text(size = 12, face = "bold"),
            axis.text = element_text(size = 12),
            strip.text.x = element_text(size = 12, face = "bold"),
            legend.title=element_text(size=12, face = "bold"),
            legend.text=element_text(size=12),
            panel.background = element_rect(fill = "azure1"))
  
```

```{r}
      #### Define color mapping for each parameter
color_mapping <- c(
        "temp" = "#FF9800",
        "PAR" = "#FFE082",
        "pressure" = "#607D8B",
        "oxygen_conc" = "#4DB6AC",
        "perc_ox_sat" = "#4DD0E1",
        "fluorescence" = "#4CAF50")
      
```



# Last clean up for FCM data and merging FCM an CTD data
```{r}
fcm <- fcm %>% 
  group_by(site, depth) %>% 
  mutate_all(mean) %>% distinct() %>% 
  mutate(site = ifelse(site == "DEEP", "Deep", site)) %>% 
  rename(station = site) %>% ungroup() %>% 
  pivot_longer(cols= c(hbac_rel_abun:pro_rel_abun),
               names_to = "microbe",
               values_to = "rel_abun") %>% 
  mutate(depth = as.numeric(depth))

CTD_FCM<- left_join(fcm, CTD, by = c("depth", "station"))
```

# Write new CSV with the combined data 
```{r}
write.csv(CTD_FCM, here("Oceanographic_data", "data", "CTD_FCM.csv"))
```


# To do the microbe distribution plots I needed all my microbes in one column
```{r}
fcm <- read_csv(here("Oceanographic_data", "data", "fcm.csv"))
```

# Putting all the microbes in one column
```{r}
fcm_long <- fcm %>% rename(
  Prochlorococcus = pro_rel_abun,
  Synechococcus = syn_rel_abun,
  `Pico Eukaryote` = pico_euk_rel_abun,
  `Heterotrophic Bacteria` = hbac_rel_abun) %>% 
  pivot_longer(cols = Prochlorococcus:`Heterotrophic Bacteria`,
               names_to = "microbes",
               values_to = "rel_abun")


```

# Writing a new file
```{r}
write_csv(fcm_long, here("Oceanographic_data", "data", "fcm_long.csv"))
```

